\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-07 papers/itsc_2012/draft2.tex
architecture diagram finalized

\usepackage[pdftex]{graphicx}
% TODO remember we need to take out all the color
\newcommand{\pix}[3]{
  \begin{figure}[h]
    \centering \includegraphics[#3]{#1}
    \caption{#2}
  \end{figure}
}

@@

\pix{architecture.png}
    {An illustration of AORTA's components.}
    {scale=0.2}

@@

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder marks GPS locations common to multiple roads, since
        they implicitly indicate intersections.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each direction
        (unless the road is marked one-way). The builder guesses the number of lanes based on
        OSM's ``road type'' tag or an explicit number of lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        vertex. The builder uses an angle heuristic to choose between left, right, straight,
        and U-turns. When several lanes all cross into fewer lanes,
        the builder forces merging of the rightmost lanes.  Although inaccurate,
        these patterns are the best compromise. At the time of writing, even sources
        like Google Maps \cite{googleMaps} lack this information.
  \item Because turns are heuristically placed, the graph is often
        disconnected. In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the digraph.
        Disconnected lanes are spliced out of larger roads, leaving unappealing
        geometric gaps, but correctly ensuring connectivity for corner-case free
        pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Simulation is discrete-time, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity, and
space-continuous, meaning roads are not tiled. The step duration $dt$ is fixed,
enabling agents to reason about collision avoidance and satisfying other
constraints. Setting this value to the average ``reaction time'' of a human
driver may be appropriate for some experiments. Each step, the simulation
performs the following:

\begin{enumerate}
  \item Introduce new agents into the map
  \item Update the position and velocity of agents based on their choice of
  acceleration in the previous step
  \item Check for collisions
  \item Allow each agent to choose a reaction for the next step
\end{enumerate}

@@

Each agent has a \emph{behavior} governing it. When the agent travels past the end
of a lane during its step, the behavior picks the turn to take. At each step,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (when the agent is at rest and is done with its
route) or setting an acceleration for the next step. Were lane changing or
merging modeled, that would be another possible action.
\commentm{What's the difference between lane changing and merging?}

@@

This paper has presented AORTA, a new city-scale traffic simulation framework
that focuses on ease-of-use and policy configurability. In conjunction with Open
Street Maps, AORTA allows researchers to repeat an experiment in any number of
places with minimal effort. Future work will exploit the flexible infrastructure
of interactions between agent behaviors and intersections by exploring dynamic
replanning and adaptable signal timings. On top of these structures,
applications could experiment with dynamic contra-flow \cite{ITSC11-hausknecht}
and micro-tolling \commentd{need a cite}. There will also be an effort to
improve AORTA as a framework by including a lane-changing model and preventing
gridlock.
\commentm{Perhaps mention concurency plans (with TACC) to simulate arbitrarily large maps?}

>>>>

\usepackage[pdftex]{graphicx}
% TODO remember we need to meet color compliance:
% http://its.papercept.net/conferences/support/general.php
\newcommand{\pix}[3]{
  \begin{figure}[h]
    \centering \includegraphics[#3]{#1}
    \caption{#2}
  \end{figure}
}

@@

% Diagrma made with gliffy.com. architecture.gxml should enable editing.
\pix{architecture.png}
    {An illustration of AORTA's components.}
    {scale=0.18}

@@

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder marks GPS locations common to multiple roads, since
        they implicitly indicate intersections.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each direction
        (unless the road is marked one-way). The builder guesses the number of lanes based on
        OSM's ``road type'' tag or an explicit number of lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        junction. The builder uses an angle heuristic to choose between left, right, straight,
        and U-turns. When several lanes all cross into fewer lanes,
        the builder forces merging of the rightmost lanes.  Although inaccurate,
        these patterns are the best compromise. At the time of writing, even sources
        like Google Maps \cite{googleMaps} lack this information.
  \item Because turns are heuristically placed, the graph is often
        disconnected. In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the digraph.
        Disconnected lanes are spliced out of larger roads, leaving unappealing
        geometric gaps, but correctly ensuring connectivity for corner-case free
        pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Simulation is discrete-time, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity, and
space-continuous, meaning roads are not tiled. The step duration $dt$ is fixed,
enabling agents to reason about collision avoidance and satisfying other
constraints. Setting this value to the average ``reaction time'' of a human
driver may be appropriate for some experiments. Each step, the simulation
does the following:

\begin{enumerate}
  \item Introduces new agents into the map
  \item Updates the position and velocity of agents based on their choice of
  acceleration in the previous step
  \item Checks for collisions
  \item Allows each agent to choose a reaction for the next step
\end{enumerate}

@@

Each agent has a \emph{behavior} governing it. When the agent travels past the end
of a lane during its step, the behavior picks the turn to take. At each step,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (when the agent is at rest and is done with its route)
or setting an acceleration for the next step. Were lane changing modeled, that
would be another possible action.

@@

This paper has presented AORTA, a new city-scale traffic simulation framework
that focuses on ease-of-use and policy configurability. In conjunction with Open
Street Maps, AORTA allows researchers to repeat an experiment in any number of
places with minimal effort. Future work will exploit the flexible infrastructure
of interactions between agent behaviors and intersections by exploring dynamic
replanning and adaptable signal timings. On top of these structures,
applications could experiment with dynamic contra-flow \cite{ITSC11-hausknecht}
and micro-tolling \commentd{Dr. Stone, should we just not mention this yet since
it's not well-published?}. There will also be an effort to improve AORTA as a
framework by including a lane-changing model and preventing gridlock.
\commentm{Perhaps mention concurency plans (with TACC) to simulate arbitrarily large maps?}
